Supplemental Materials for ``Universals of word order result from optimization of grammars for efficient communication''

### 2019-09-12 writeup/pnas-word-order-si.tex
code
SI

@@

>>>>

\section{Other Methods of Estimating Efficiency and Efficiency Optimization}

\subsection{Lexicalized Models}

\subsection{Original UD Format}

\subsection{Nondeterministic Baseline Grammars}

@@

\section{Simulation for Parseability and Correlating Orders}

%VP -> V
%VP -> VP[head] VP[xcomp]
%VP -> VP[head] NP[obj]
%
%NP -> N
%NP -> NP[head] VP[acl]

We consider the following head-outward generative model.
Verbs generate verb dependents (\emph{xcomp}) and noun dependents (\emph{obj}), independently.
The overall number $N$ of dependents is $NB(1, p_{branching})$, the number of \emph{obj} dependents is $Binom(p_{obj}, N)$.
Nouns can generate verb dependents (\emph{acl}), of number $NB(1, p_{acl})$.

Trees are linearized using one of two grammars: One (`Correlating') places \emph{obj}, \emph{xcomp}, and \emph{acl} dependents on the same side of the head, and (in accordance with crosslinguistic tendencies, which are predicted by DLM) places the \emph{obj} dependents closer to the head than \emph{xcomp} dependents.
The other grammar (`Anti-Correlating') places \emph{xcomp} and \emph{acl} dependents opposite to \emph{obj} dependents.

We formulated this model as a binary-branching PCFG, and used a CKY parser to estimate $I[\mathcal{T}, \mathcal{U}]$ from 10,000 random sample sentences.

We computed this for different settings of $p_{branching} \in [0, 0.5]$ and $p_{obj} \in [0,1]$, at $p_{acl} \in \{0, 0.3\}$.\footnote{For $p_{acl} = 0.3$, we only computed values for $p_{branching} \leq 0.4$, due to high computational cost on long sentences resulting from $p_{branching} = 0.5$ and $p_{acl} = 0.3$.}
For these settings, we computed the difference in $I[\mathcal{T}, \mathcal{U}]$ between the two grammars.
Results are shown in Figure~\ref{fig:toy-parseability}.

\begin{figure}
	\begin{center}	
\includegraphics[width=0.4\textwidth]{../models/revision/toy_simulations/result_NPBranching0_0.pdf}
\includegraphics[width=0.4\textwidth]{../models/revision/toy_simulations/result_NPBranching0_3.pdf}
\end{center}

	\caption{Difference in parseability $I[\mathcal{T}, \mathcal{U}]$ (normalized by sentence length) between the correlating and anti-correlating grammars, for $p_{acl} = 0.0$ (left) and $p_{acl} = 0.3$ (right). Positive values indicate greater values of $I[\mathcal{T}, \mathcal{U}]$ for correlating grammars.}\label{fig:toy-parseability}
\end{figure}
